MS Technical Paper: \\ Placement Algorithms for Heterogeneous FPGAs

### 2025-04-22 latex/paper/paper.tex
minor edit
minor edits

    Field-Programmable Gate Arrays (FPGAs) have witnessed rapid growth in capacity and versatility, driving significant advances in computer-aided design (CAD) and electronic design automation (EDA) methodologies. 
    Since the early-to-mid 2000s, the stagnation of single-processor performance relative to the rapid increase in integrated circuit sizes has led to a design productivity gap, where the computational effort for designing complex chips continues to rise. 
    FPGA CAD flows mainly encompass synthesis, placement, and routing; all of which are HP-hard problems, of which placement is one of the most time-consuming processes. 
    Inefficienct placement strategy not only extends design times from hours to days, thereby elevating cost and reducing engineering productivity, but also limits the broader adoption of FPGAs by software engineers who expect compile times akin to those of conventional software compilers like {\tt gcc}.

@@

    \textbf{FFs} \quad
        FFs are synthesized to facilitate synchronous event-driven signal assignment. 
        For most Verilog users, this generally means signal assignments wrapped in \texttt{always @(posedge clk)} statements. 
        Figure \ref{fig:ff_synthesis} shows an example of where a flip flop (FF) is typically synthesized. 
        The cell primitive FDRE is a type of FF and belongs to a family of D Flip Flops (DFFs) with Clock Enable (CE). 
        \begin{itemize}
            \item \emph{FDCE - DFF with CE and Asynchronous Clear}
            \item \emph{FDPE - DFF with CE and Asynchronous Preset}
            \item \emph{FDSE - DFF with CE and Synchronous Set}
            \item \emph{FDRE - DFF with CE and Synchronous Reset}
        \end{itemize}
        In a typical HDL design, the vast majority of FFs will be synthesized as FDREs with the occasional FDSE, as it is generally good practice to keep FPGA designs synchronous. 
        A FF BEL may also host a latch cell, however, since they are generally bad practice in FPGA design, we will not consider latches in this paper. 
        {
            \centering
            \includegraphics[width=\columnwidth]{figures/ff_synthesis.png}
            \captionof{figure}{FF synthesis from user design}
            \label{fig:ff_synthesis}
        }

@@

    \textbf{CARRY Chains} \quad
        These CARRY4 blocks can be chained across SLICEs to implement wide adders efficiently. 
        The CARRY4 BELs \emph{must} be chained vertically consecutively across SLICEs as the Carry-In (CI) and Carry-Out (CO) pins can only be routed this way. 
        These CARRY4 BELs may also be directly connected to LUTs and FFs, and should be placed in the same Site whenever possible to minimize wirelength. 
        Shown in figure \ref{fig:carry_cell_group} shows how a CARRY4 chain and associated LUTs and FFs can be placed across SLICEs.

@@

    \textbf{DSP Cascades} \quad 
        Higher level DSP functions are supported by cascading DSP48E1 slices in a DSP48E1 column.
        Much like CARRY4 chains, DSP48E1 cascades must necessarily be placed vertically consecutively across DSP48E1 Sites. 
        They are connected by three 48-bit signal busses: ACOUT to ACIN, BCOUT to BCIN, PCOUT to PCIN.
        These signal busses run directly between the vertical DSP48E1 slices without burdening the general routing resources. 
        The ability to cascade this way provides a high-performance and low-power implementation of DSP filter functions.

>>>>

    Field-Programmable Gate Arrays (FPGAs) have witnessed rapid growth in capacity and versatility, driving significant advances in computer-aided design (CAD) and electronic design automation (EDA) methodologies. 
    Since the early-to-mid 2000s, the stagnation of single-processor performance relative to the rapid increase in integrated circuit sizes has led to a design productivity gap, where the computational effort for designing complex chips continues to rise. 
    FPGA CAD flows mainly encompass synthesis, placement, and routing; all of which are NP-hard problems, of which placement is one of the most time-consuming processes. 
    Inefficienct placement strategy not only extends design times from hours to days, thereby elevating cost and reducing engineering productivity, but also limits the broader adoption of FPGAs by software engineers who expect compile times akin to those of conventional software compilers like {\tt gcc}.

@@

    \textbf{FFs} \quad
        FFs are synthesized to facilitate synchronous event-driven signal assignment. 
        For most Verilog users, this generally means signal assignments wrapped in \texttt{always @(posedge clk)} statements. 
        Figure \ref{fig:ff_synthesis} shows an example of where a flip flop (FF) is typically synthesized. 
        The cell primitive FDRE is a type of FF and belongs to a family of D Flip Flops (DFFs) with Clock Enable (CE). 
        \begin{itemize}
            \item \emph{FDCE - DFF with CE and Asynchronous Clear}
            \item \emph{FDPE - DFF with CE and Asynchronous Preset}
            \item \emph{FDSE - DFF with CE and Synchronous Set}
            \item \emph{FDRE - DFF with CE and Synchronous Reset}
        \end{itemize}
        In a typical HDL design, the vast majority of FFs will be synthesized as FDREs with the occasional FDSE, as it is generally good practice to keep FPGA designs synchronous. 
        A FF BEL may also host a LATCH Cell, however, since they are generally bad practice in FPGA design, we will not consider latches in this paper.

        {
            \centering
            \includegraphics[width=\columnwidth]{figures/ff_synthesis.png}
            \captionof{figure}{FF synthesis from user design}
            \label{fig:ff_synthesis}
        }

@@

    \textbf{CARRY Chains} \quad
        These CARRY4 blocks can be chained across SLICEs to implement wide adders efficiently. 
        The CARRY4 BELs \emph{must} be chained vertically consecutively across SLICEs as the Carry-In (CI) and Carry-Out (CO) pins can only be routed this way. 
        A CARRY4 cell may also directly connect to LUTs and FFs, and should be placed in the same Site whenever possible to minimize wirelength. 
        Shown in figure \ref{fig:carry_cell_group} shows how a CARRY4 chain and associated LUTs and FFs can be placed across SLICEs.

@@

    \textbf{DSP Cascades} \quad 
        Higher level DSP functions are supported by cascading DSP48E1 slices in a DSP48E1 column.
        Much like CARRY4 chains, DSP48E1 cascades must necessarily be placed vertically consecutively across DSP48E1 Sites. 
        They are connected by three busses: ACOUT to ACIN, BCOUT to BCIN, PCOUT to PCIN.
        These signal busses run directly between the vertical DSP48E1 slices without burdening the general routing resources. 
        The ability to cascade this way provides a high-performance and low-power implementation of DSP filter functions.
